Regressions

### 2020-01-22 analysis/reports/regressions.Rmd
more regressions and plots

  write.csv(all_di, "all_di.csv", row.names = F)
} else {
  all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
}
```

```{r histograms, fig.height = 6, fig.width =6}

skewhists <- ggplot(data = all_di, aes(x = skew_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3, boundary = 0) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500), boundary = 0) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Skewness") +
  geom_vline(xintercept = 95, color = "red")
skewhists

simpsonhists <- ggplot(data = all_di, aes(x = simpson_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3, boundary = 0) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500), boundary = 0) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Simpson") +
  geom_vline(xintercept = 5, color = "red")

simpsonhists

@@

Observations from these histograms:

* FIA really struggles to get 2500 samples
* Portal is the *most* concentrated of all the datasets.
* Then MCDB and misc, then gentry, then bbs. FIA is hard to place but looks idiosyncratic.
* Curious about how variation in %ile maps on to S and N, *particularly* for Gentry and BBS. 
* Skewness is less extreme than Simpson (more density in the in-between). 
* Singletons doesn't appear to be terrifically impactful. From now on filtering to *unaltered* vectors.
* I think it's best to filter to at least 2000 samples.

Some summary statistics:

```{r summary plots}

di_summary <- all_di %>%
  filter(nsamples >= 2000) %>%
  mutate(avg_n = n0 / s0,
         skew_pp = skew_percentile / 100,
         simp_pp = simpson_percentile/100)
```

### Avg abundance

```{r avg abundance regression}

avgn_skew <- ggplot(data = di_summary, aes(x = avg_n, y = skew_pp)) +
  geom_point() +
  stat_smooth(method = "glm",method.args = list(family = binomial)) + 
  facet_wrap(vars(dat), scales = "free") +
  theme_bw()

avgn_skew

summary(glm(di_summary, formula = skew_pp ~ avg_n * dat, family = binomial))

for(i in unique(di_summary$dat)) {
  this_dat <- filter(di_summary, dat==i)
  print(i)
  print(summary(glm(skew_pp ~ avg_n, family = binomial,
          data = this_dat)))
}

@@

>>>>

  write.csv(all_di, "all_di.csv", row.names = F)
} else {
  all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
}

di_summary <- all_di %>%
  filter(nsamples >= 2000, !singletons) %>%
  mutate(avg_n = n0 / s0,
         skew_pp = skew_percentile / 100.01,
         simp_pp = simpson_percentile/100.01)

@@

Filtering to nsamples >= 2000, not singletons.

### Supposed correlates

#### Average abundance (N/S)

```{r avg n}

avg_n_plot <- ggplot(data = di_summary, aes(x = avg_n)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free") +
  theme_bw() +
  ggtitle("Average abundance")
print(avg_n_plot)

avg_n_boxplot <- ggplot(data = di_summary, aes(x= dat, y = avg_n)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Average abundance")

print(avg_n_boxplot)

avg_n_boxplot_log <- ggplot(data = di_summary, aes(x= dat, y = log(avg_n))) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Average abundance")

print(avg_n_boxplot_log)
```

```{r means and sds}
skew_mean_plot <- ggplot(data = di_summary, aes(x = dat, y = skew_mean)) +
  geom_boxplot() +
  theme_bw()
skew_sd_plot <- ggplot(data = di_summary, aes(x = dat, y = skew_sd)) +
  geom_boxplot() +
  theme_bw()
skew_m_sd_plot <- ggplot(data = di_summary, aes(x = skew_mean, y = skew_sd, color = dat)) +
  geom_point(alpha = .2) +
  theme_bw()

skew_mean_plot
skew_sd_plot
skew_m_sd_plot

simpson_mean_plot <- ggplot(data = di_summary, aes(x = dat, y = simpson_mean)) +
  geom_boxplot() +
  theme_bw()
simpson_sd_plot <- ggplot(data = di_summary, aes(x = dat, y = simpson_sd)) +
  geom_boxplot() +
  theme_bw()
simpson_m_sd_plot <- ggplot(data = di_summary, aes(x = simpson_mean, y = simpson_sd, color = dat)) +
  geom_point(alpha = .2) +
  theme_bw()

simpson_mean_plot
simpson_sd_plot
simpson_m_sd_plot

@@

The different datasets are on dramatically different scales of average abundance, although it looks like a lot of that is outliers in misc_abund.

### Correlation plots

#### Average abundance
```{r avg n vis, fig.dim = c(8,8)}

skew_avgn_plot <- ggplot(data = di_summary, aes(x = avg_n, y = skew_percentile)) +
  geom_point(size = 1, alpha = .2) +
  facet_wrap(vars(dat), scales = "free_x") +
  theme_bw() +
  ggtitle("Skew v Avg N")
print(skew_avgn_plot)

simpson_avgn_plot <- ggplot(data = di_summary, aes(x = avg_n, y = simpson_percentile)) +
  geom_point(size = 1, alpha = .2) +
  facet_wrap(vars(dat), scales = "free_x") +
  theme_bw() +
  ggtitle("Simpson v Avg N")
print(simpson_avgn_plot)

#### Mean and sd

```{r mean sd vis, fix.dim = c(8, 8)}

skew_mean_plot <- ggplot(data = di_summary, aes(x = skew_mean, y = skew_percentile)) +
  geom_point(size = 1, alpha = .2) +
  facet_wrap(vars(dat), scales = "free") +
  theme_bw() +
  ggtitle("Skew v Mean skew")
print(skew_mean_plot)

skew_sd_plot <- ggplot(data = di_summary, aes(x = skew_sd, y = skew_percentile)) +
  geom_point(size = 1, alpha = .2) +
  facet_wrap(vars(dat), scales = "free") +
  theme_bw() +
  ggtitle("Skew v sd skew")
print(skew_sd_plot)

simpson_mean_plot <- ggplot(data = di_summary, aes(x = simpson_mean, y = simpson_percentile)) +
  geom_point(size = 1, alpha = .2) +
  facet_wrap(vars(dat), scales = "free") +
  theme_bw() +
  ggtitle("simpson v Mean simpson")
print(simpson_mean_plot)

simpson_sd_plot <- ggplot(data = di_summary, aes(x = simpson_sd, y = simpson_percentile)) +
  geom_point(size = 1, alpha = .2) +
  facet_wrap(vars(dat), scales = "free") +
  theme_bw() +
  ggtitle("simpson v sd simpson")
print(simpson_sd_plot)

### Beta regression

#### Average abundance

```{r betareg avg n}
library(betareg)

di_summary$skew_pp[ which(di_summary$skew_pp == 0)] <- .00000001
di_summary$simp_pp[ which(di_summary$simp_pp == 0)] <- .00000001

avgn_br_skew <- betareg(skew_pp ~ avg_n + dat, data = di_summary)
summary(avgn_br_skew)
plot(avgn_br_skew)
